---
title: "Quick Analysis"
author: "ahs"
date: "November 29, 2015"
output: pdf_document
---

This is a quick little analysis I did for some data that I collected as part 
of my final dissertation project. I had about 20 people use a food diary on their 
cell phones for 3 weeks. The code was instrumented with analytics codes to make it 
easy (well, possible!) to analyze usage of the food diary over time. 

```{r, echo = FALSE, warning = FALSE}

data_file = '~/Documents/data/POND_actionlogs/PondActionLog.csv'

action_map_file =  '~/Documents/data/POND_actionlogs/PondActionIdMapping.csv'

histo_colors = c('darkorange1', 
                 'darkorchid', 
                 'deepskyblue1', 
                 'gold1', 
                 'deeppink1', 
                 'yellowgreen', 
                 'cadetblue', 
                 'maroon1', 
                 'navy', 
                 'tan4', 
                 'black')


  library(gdata)
  library(ggplot2)
  library(plyr)
  library(dplyr)
  library(class)
  library(gmodels)


action_log <- read.csv(data_file)
head(action_log)

actionId_map <- read.csv(action_map_file)

nrow(action_log)
action_log <- action_log[action_log$pptId != '',]
action_log$when <- strptime(action_log$timestamp, format='%m/%d/%y %H:%M')
action_log$when.l <- as.integer(as.POSIXct(action_log$when))

## Fine for a moment, but should really be shifted by time of day when the study started
action_log$studyDay <- floor(action_log$mins / 1440)

action_log <- merge(action_log, actionId_map, by = 'actionId', all.x = TRUE )
action_log <- unique(action_log) ## It looks like there are 3-4 copies of each log in the data set. 

# actionId == 14 => start of study

table(action_log$typeId)

p1001 <- action_log[action_log$pptId == 'p1001', ]
p1001$when <- strptime(p1001$timestamp, format='%m/%d/%y %H:%M')
p1001$when.l <- as.integer(as.POSIXct(p1001$when))
p1001.start <- p1001[p1001$actionId == 14, ]

```

My first plot is all the data, with each action type represented in each facet panel. 
This first one is just the action for a single participant on a single day, so a 
point represents that the given participant had a given action on that day of the study. 


```{r, echo=FALSE, fig.height = 8}

minToDayLabels <- lapply(seq(0, 42373, by = 60 * 24), 
                         FUN = function(x)(return(sprintf('Day %d', x/1440))))


ggplot(data = unique(action_log), aes(x = studyDay, y = pptId, 
                                      fill = actionName, 
                                      color = actionName)) + 
  scale_color_manual( values = histo_colors) +
  scale_x_continuous(breaks = seq(0, 29), 
                     labels= minToDayLabels) +
  #geom_jitter(binwidth = 60 ) + 
    geom_vline(xintercept = 0) + 
  geom_vline(xintercept =  7) + 
  geom_vline(xintercept =  14) + 
  geom_vline(xintercept =  21) + 
  theme(axis.text.x = element_text(angle=90, vjust = 1, hjust = 1), 
        axis.text.y = element_blank()) +
  geom_point() + 
  facet_grid(actionName ~ .) + 
  ggtitle('All Ppts, All Actions')


```

A few initial observations: 

* Only a couple of participants changed any daily goals (purple dots), and that was only in 
  the first few days of the study. 
* More people actually looked at the goals (black dots), and that was throughout the study. 
* A few people edited records throughout the study, but not many. 

There are three things we want to look at: 

* Actions related to using the food database. 
* Actions related to using the +1 buttons. 
* How often the app was started over the course of the study. 

# App Starts throughout the study

One behavior we wanted to see from our participants was a good pattern of 
starting the app-- multiple times per day, hopefully a consistent pattern. 
Here, I've plotted the number of app starts per day. The black horizontal line 
indicates three app starts per day. 

```{r, echo = FALSE, warning = FALSE, fig.height = 12}

starts.only <- action_log[action_log$actionId == 0,
                           c('pptId', 'studyDay', 'actionId')]
num.starts.per.day <- ddply(starts.only, 
                           c('pptId', 'studyDay'), count)

num.starts.per.day.plot <- 
  ggplot(data = num.starts.per.day, aes(x = studyDay, y = n, 
                                      colour = pptId)) + 
  #scale_color_brewer(type = 'qual', palette = 'Set1') +
  scale_x_continuous(breaks = seq(-1, 29, by = 1), 
                     labels= c('Day -1', minToDayLabels)) +
  theme_light() +
  geom_vline(xintercept = 0) + 
  geom_vline(xintercept =  7) + 
  geom_vline(xintercept = 14) + 
  geom_vline(xintercept = 21) + 
  geom_hline(yintercept = 3) + 
  theme(axis.text.x = element_text(angle=90, vjust = 1, hjust = 1)) +
  geom_line(size = 1) + 
  facet_grid(pptId ~ .) + 
  ggtitle('Number of times app was started each day')

print(num.starts.per.day.plot)

```


* to do: table of number of app starts per day. Mean number of starts per day for each ppt. 
* For each ppt: 
    * Number of days with more than one start
    * Number of days with 0 starts 
    * Number of days with 3 or more starts
    
# Question to investigate: 

In the preliminary analysis, I saw different users use different strategies for 
entering given food items. It would be interesting to see if there is a correlation 
between what people do in the lab and what they do in the field. 



